The Efficient Handling of Guards \\in the Design of RPython's Tracing JIT

### 2012-09-06 talk/vmil2012/paper.tex
done
cleanup
tweak colors
describe second kotzmann paper

\usepackage{ifthen}
\usepackage{sparklines}
\usepackage{booktabs}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage[pdfpagelabels=true]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}

@@

\definecolor{commentgray}{rgb}{0.3,0.3,0.3}

@@

\bigtodo{reviewer c: It might be possible to use a chart for
Figure~\ref{fig:failing_guards} to give more information?}

@@

The deoptimization information of Hotspot is extended
to support correct behaviour
when scalar replacement of fields is done for non-escaping objects~\cite{kotzmann_escape_2005}.
The approach is extremely similar to how RPython's JIT handles virtual objects.
For every object that is not allocated in the code,
the deoptimization information contains a description
of the content of the fields.
When deoptimizing code, these objects are reallocated
and their fields filled with the values
described by the deoptimization information.
The paper does not describe any attempts to store this information compactly.
This may not be needed in their approach, because method-based JITs have
fewer deoptimization points than tracing JITs.

\bigtodo{reviewer d: You state that Kotzmann et al. do not explain how the
debug information is store in a compact way. They have a second paper that
describes exactly that: Thomas Kotzmann, Hanspeter Mössenböck: Run-Time
Support for Optimizations Based on Escape Analysis Proceedings of the
International Symposium on Code Generation and Optimization, San Jose,
March 2007.}

>>>>

\usepackage{ifthen}
\usepackage{sparklines}
\usepackage{booktabs}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{wrapfig}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage[pdfpagelabels=true]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}

@@

\definecolor{commentgray}{rgb}{0.3,0.3,0.3}
\definecolor{darkgray}{rgb}{0.5,0.5,0.5}
\definecolor{lightgray}{rgb}{0.75,0.75,0.75}

@@

@@

The deoptimization information of Hotspot is extended
to support correct behaviour
when scalar replacement of fields is done for non-escaping objects~\cite{kotzmann_escape_2005, kotzmann_run-time_2007}.
The approach is extremely similar to how RPython's JIT handles virtual objects.
For every object that is not allocated in the code,
the deoptimization information contains a description
of the content of the fields.
When deoptimizing code, these objects are reallocated
and their fields filled with the values
described by the deoptimization information.
The data structures for the deoptimization information are very similar to
those used by RPython's tracing JIT. For every compiled Java method there is a
\emph{scope entry} for the stack and one for the local variables. The objects
that are replaced by scalars are described by \emph{object entries}, which are
equivalent to RPython's virtual objects.

The papers does not describe any attempts to share the object entries and scope
entries between different deoptimization safe points. This seems to not be
needed in a method-based JIT compiler, because method-based JITs have fewer
deoptimization points than tracing JITs. Indeed, in the evaluation presented in
the second paper~\cite{kotzmann_run-time_2007} the number of safe points is low
for the benchmarks presented there, between 167 and 1512. The size of the
debugging information in the presented benchmarks is at most about half the
size of the machine code generated.

\bigtodo{for comparison we should add the total number of guards per benchmark to one of the tables}
